---
title: "Replication"
format: pdf
author: "Peng-Ting Kuo"
---

# Introduction

This file is the replication for the paper:

Meijers, M. J., & Zaslove, A. (2021). Measuring Populism in Political Parties: Appraisal of a New Approach. *Comparative Political Studies*, *54*(2), 372-407.

# Setup

```{r Package and Data}
library(ggplot2)
library(tidyverse)
library(haven)
library(psych)

# Expert level data
expert <- read_stata("replication/data/expert_data_stata.dta")

# Expert mean data (these two datasets are almost identical)
expert_mean <- read_stata("replication/data/party_means.dta")
expert_mean <- readRDS("replication/final_data/poppa1.rds")
```

# Data Processing

The expert level dataset I have is different from the original one. (Missing peopleimm, peoplewealthy, elitepol, elitemedia, eliteecon, climatepolicy, climatescience, democracy, positionchange, compromise

```{asis Variable Explanation}
Note 1: The Sweden has been asked at a 1-10 scale.  

Variables:
complex: Some parties portray political decision-making as a complex process, while others
advocate common-sense solutions to political problems.
0 = Not at all complex; 10 = Highly complex

emotional:  Some parties appeal to emotions in their political communication with the voter
(i.e fear, hope, anger, happiness).
0 = Not at all; 10 = Very much

# Party Organization
intradem: Some political parties practice more intra-party democracy than others (i.e. party
members play a role in decision making, there is room for internal debate, decision-making is
inclusive of various factions and organizational layers within the party).
0 = No intra-party democracy at all; 10 = Very high level of intra-party democracy

personalised: Parties can be characterized by more or less personal leadership.
Please tick the box that best describes each party's degree of personalized leadership.
0 = Not at all personalized; 10 = Very personalized

# Party Ideology
lroverall: Please tick the box that best describes each party's overall ideology on a scale
ranging from 0 (left) to 10 (right) (i.e. the general left-right scale).
0 = Left; 10 = Right

lrecon: Parties can be classified in terms of their stance on economic issues. Parties on the
economic left want government to play an active role in the economy. Parties on the economic
right emphasize a reduced economic role for government: privatization, lower taxes, less
regulation, less government spending, and a leaner welfare state.
0 = Economic left; 10 = Economic right

immigration: Some parties are strongly in favour of immigration, while others are strongly
opposed to immigration.
0 = Strongly opposed to immigration; 10 = Strongly in favor of immigration

eu: Some parties are strongly in favour of the EU and European integration, while others are
strongly opposed to the EU and European integration.
0 = Strongly opposed to the EU; 10 = Strongly in favor of the EU

nativism: Some parties have an exclusive idea of who can and should belong to the nation-
state (nativism).
Please tick the box that best describes the extent to which each party regards non-native
elements (persons or ideas) in society as a threat to the nation’s culture, well-being or
sovereignty.
0 = Not at all nativist; 10 = Very nativist

laworder: Parties can be classified in terms of their stance regarding civil liberties (i.e.
personal freedoms) and law and order (i.e. stricter policing, harsher sentencing).
Please tick the box that best describes the extent to which each party prioritizes civil liberties or law and order.
0 = Strongly promotes civil liberties; 10 = Strongly supports law and order

lifestyle: Parties can be classified in terms of their stance regarding social lifestyle. Some hold a traditional view on moral values (such as conventional notions of the family: i.e. marriage and child-rearing), while others hold more liberal views on moral values (i.e. acceptance of homosexuality, same-sex marriage etc.).
0 = More traditional; 10 = More liberal

# Expert-level question
gender: 0: male; 1: female, 2: other

self_leftright: Where would you place yourself on the general left-right scale?
0: left, 10: right

discipline: Please designate your primary academic discipline.
1 = Political Science.  
2 = Public Administration (Public Management).  
3 = Communication Science.  
4 = Sociology.  
5 = Law and Legal Studies.  
6 = History.  
7 = Other.  

expertise:  Please indicate your research expertise (multiple answers are possible).
1 = Party Politics.  
2 = Populism Studies.  
3 = Radical Right / Left and/or Political Extremism.  
4 = Voting Behavior. 
5 = Area Studies.  
6 = Public Policy.  
7 = EU Politics.  
8 = Interest Groups.  
9 = Political Economy.  
10 = Social Movements and Protest Politics.  
11 = International Relations.  
12 = Political Theory.  
13 = Political Communication.  
14 = Law and Politics.  
15 = Other.  
```

```{r Data Preprocessing}
# We only have the clean data 
# 1. Follow the authors' procedure and rename the wrong column
expert_dataset <- expert %>%
  rename(response_id = expert_id,
         indivisible = indivisble) 

columns <- c("manichean", "indivisible", "generalwill", "peoplecentrism", "antielitism", "lroverall", "lrecon", "immigration", "nativism", "lifestyle", "laworder", "laworder", "eu", "intradem", "personalised", "complex", "emotional")

# 2. reorder the columns
expert_dataset <- expert_dataset %>%
  select(
    response_id, country_id, nr_experts, party, party_id,
    all_of(columns),                                     
    everything()                                         
  )
```

## Expert-level to Party-level

```{r}
party_level <- expert_dataset %>%
  group_by(country_id, party_id) %>%
  summarise(
    n_experts = n(),
    across(manichean:emotional, 
           ~ if(sum(!is.na(.x)) >= 4) mean(.x, na.rm = TRUE) else NA_real_),
    .groups = "drop"
  ) %>%
  mutate(
    pop_vars = list(c("manichean", "indivisible", "generalwill", "peoplecentrism", "antielitism")),
    valid_count = rowSums(!is.na(select(., all_of(pop_vars[[1]])))),
    populism = ifelse(valid_count >= 4, 
                      rowMeans(select(., all_of(pop_vars[[1]])), na.rm = TRUE), 
                      NA_real_)
  ) %>%
  select(-pop_vars, -valid_count) %>%
  relocate(populism, .before = manichean)
```

```{r Data Aggregation}
# Load the original data
fa_df <- read_stata("./replication/Paper/dta_1.dta")

comparison_df <- data.frame(
  id = 1:nrow(expert_mean),
  manichean_diff   = party_level$manichean - fa_df$manichean,
  indivisble_diff  = party_level$indivisible - fa_df$indivisble,
  generalwill_diff = party_level$generalwill - fa_df$generalwill,
  peoplecentrism_diff = party_level$peoplecentrism - fa_df$peoplecentrism,
  antielitism_diff = party_level$antielitism - fa_df$antielitism
)

comparison_long <- comparison_df %>%
  pivot_longer(cols = ends_with("_diff"), 
               names_to = "variable", 
               values_to = "difference")

# Visualization
ggplot(comparison_long, aes(x = difference)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free") + 
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", alpha = 0.5) +
  labs(
    title = "Stata vs R",
    x = "Expert file - DTA file",
    y = "Frequency"
  ) +
  theme_minimal()

summary_stats <- comparison_long %>%
  group_by(variable) %>%
  summarise(
    mean_diff = mean(difference, na.rm = TRUE),
    max_diff  = max(abs(difference), na.rm = TRUE),
    sd_diff   = sd(difference, na.rm = TRUE)
  )

print(summary_stats)
```

# Factor Analysis

Note that the authors use Stata for the table, the result in R is slightly different (around 0.01-0.02).

## Table 2

```{r FA}
# My version of factor analysis
fa_df <- expert_mean %>%
  select(manichean, indivisble, generalwill, peoplecentrism, antielitism) %>%
  na.omit()
fit_fa <- factanal(x = fa_df,
                   factors = 1,
                   scores = "regression",
                   rotation = "varimax") # This is the default method

# print(fit_fa)


# Author's version of factor analysis (I convert to R version)
efa_vars_author <- fa_df %>%
  select(manichean, indivisble, generalwill, peoplecentrism, antielitism) %>%
  na.omit() 

# The dataset I transfer from expert level
efa_vars <- party_level %>%
  select(manichean, indivisible, generalwill, peoplecentrism, antielitism) %>%
  na.omit() 

# Iterated Principal Factor (IPF)
## Stata code:
# use "dta_1.dta", clear

# 	factor  manichean indivisble generalwill peoplecentrism antielitism   , ipf mineigen(1) 
# 	rotate,  orthogonal varimax

# *Create predicted populism variable
# 	predict populism_factor	
# 	gen populism = ((populism_factor - (-1.616698))/ 3.682342) * 10


# Stata's ipf' equals to psych's fm = "pa" (Principal Axis)
efa_result <- fa(efa_vars, 
                 nfactors = 1,      
                 fm = "pa",          
                 max.iter = 50,     
                 rotate = "varimax")

# factor.loading
loading <- round(efa_result$loadings, 3)
print(loading)

# uniqueness
uniqueness <- round(efa_result$uniquenesses, 3)
print(uniqueness)

# factor.scores
scores <- factor.scores(efa_vars, efa_result)$scores
expert_dataset$populism_factor <- NA


# expert_dataset$populism_factor[complete.cases(expert_dataset %>% 
#                                select(manichean, indivisible, generalwill, peoplecentrism, antielitism))] <- scores[,1]

# 5. regularization ((score - minimum) / distance) * 10
expert_dataset <- expert_dataset %>%
  mutate(populism = ((populism_factor - (-1.616698)) / 3.682342) * 10)
```

## Table 4

The authors' original stata code:

```         
factor intradem personal manichean indivisble generalwill peoplecentrism antielitism complex emotional lrecon immigration eu nativism laworder lifestyle , ipf 
rotate, blanks(.5) orthogonal varimax
```

```{r FA-2}
# 1. Dataset 
efa_vars_full <- party_level %>%
  select(
    manichean, indivisible, generalwill, peoplecentrism, antielitism, 
    lrecon, immigration, eu, nativism, laworder, lifestyle,
    complex, emotional, intradem, personalised
  ) %>%
  na.omit()

# mapping label and variable
name_mapping <- c(
  "manichean" = "Manichean worldview",
  "indivisible" = "Indivisible people",
  "generalwill" = "General will",
  "peoplecentrism" = "People-centrism",
  "antielitism" = "Anti-elitism",
  "lrecon" = "Left-Right (Economy)",
  "immigration" = "Immigration",
  "eu" = "European integration",
  "nativism" = "Nativism",
  "laworder" = "Civil liberties versus Law & order",
  "lifestyle" = "Lifestyle: Traditional versus Liberal",
  "complex" = "Complex versus Common-sense politics",
  "emotional" = "Emotional appeal",
  "intradem" = "Intra-party democracy",
  "personalised" = "Personalized leadership"
)

# 1. Eigenvalue & cor matrix
cor_matrix <- cor(efa_vars_full)
raw_eigenvalues <- eigen(cor_matrix)$values

# number of eigenvalue > 1
n_to_extract <- sum(raw_eigenvalues > 1)

# 2. Similar to Stata ipf approach
efa_optimized <- fa(
  cor_matrix, 
  nfactors = n_to_extract, 
  fm = "pa", 
  rotate = "varimax", 
  max.iter = 50,
  SMC = TRUE
)


loadings_df <- as.data.frame(unclass(efa_optimized$loadings))

# mapping names
loadings_df$Variable_Label <- name_mapping[rownames(loadings_df)]


final_paper_table <- loadings_df %>%
  transmute(
    Variable = Variable_Label,
    `Factor I` = ifelse(abs(PA1) < 0.5, "", format(round(PA1, 6), nsmall = 6)),
    `Factor II` = ifelse(abs(PA2) < 0.5, "", format(round(PA2, 6), nsmall = 6))
  )

# 5. output
cat("Factor I  Initial Eigenvalue:", round(raw_eigenvalues[1], 2), "\n")
cat("Factor II Initial Eigenvalue:", round(raw_eigenvalues[2], 2), "\n\n")
print(final_paper_table, row.names = FALSE, quote = FALSE)
```
